OpenMPIR

### 2017-10-08 paper.tex
Fixed up discussion of disabling tapir opts

It is worth noting again that the scope of the work that this paper is
responsible for is only the front end. All performance gains attributable to
the existing Tapir back end and Cilk runtime represent prior work.  We can
break reasons for performance discrepancies into a few categories:
\begin{itemize}
\item Front-end code generation;
\item IR Optimizations; and
\item Runtime efficiency.
\end{itemize}
While we would like to distinguish between these, it is difficult to do so with
only profiling information. A more thorough investigation would require deep
knowledge of each of the front ends and runtimes. While the importance of
optimizations of parallel code is clear, we hypothesize that most of the
disparity present in these results is more a function of runtime implementation
differences. As justification for this hypothesis, consider the
\texttt{Fibonacci} benchmark. There are no memory accesses to the heap, and the
code leaves little room for optimization (algorithmic rewrites aside), so it is
effectively a benchmark of how fast a runtime can execute fork-join
parallelism.

@@

There are two cases we examine using the profiler. The first is the
case when there is a large difference in run times between implementations. For 
these our hypothesis is that the slower implementations
are for some reason spending more time in the runtime. As mentioned, 
it is difficult to understand exactly why an implementation is spending more 
time in the runtime without further investigation, but it is a useful 
sanity check nonetheless. For this case, we turn to the \texttt{Fib} benchmark.
Indeed, we get roughly the following breakdowns of work-to-overhead ratio
according to the Linux \texttt{perf} tool.
\begin{itemize}
\item \texttt{tapir: ~30\% runtime overhead, ~50\% work, ~20\% other}
\item \texttt{clang: ~85\% runtime overhead, ~5\% work, ~10\% other}
\item \texttt{icc: ~85\% runtime overhead, ~5\% work, ~10\% other}
\item \texttt{gcc: ~100\% runtime overhead, ~0\% work, ~0\% other}
\end{itemize}
This supports our hypothesis that most of the slowdown incurred is overhead in
the runtime. For Tapir the overhead is relatively small given the size of the
work chunks, for \texttt{clang} and \texttt{icc} it is significantly larger,
and for \texttt{gcc} it is overwhelming. It's worth noting that \texttt{clang} 
and \texttt{icc} have effectively the same runtime, so the fact that their 
overheads are similar is unsurprising.

@@

Another way we can discriminate the cause of performance discrepency is by disabling 
Tapir optimizations. This allows us to directly measure the impact that Tapir
optimizations are having on runtime. For \texttt{fibonacci}, for example,
disabling optimizations results in a roughly ~2x performance loss. This is likely

>>>>

It is worth noting again that the scope of the work that this paper is
responsible for is only the front end. All performance gains attributable to
the existing Tapir back end and Cilk runtime represent prior work.  We can
break reasons for performance discrepancies into a few categories:
\begin{itemize}
\item Front-end code generation;
\item IR Optimizations; and
\item Runtime efficiency.
\end{itemize}
While we would like to distinguish between these, it is difficult to do so with
only profiling information. A more thorough investigation would require deep
knowledge of each of the front ends and runtimes. We hypothesize that much of
the disparity present in these results is more a function of runtime
implementation differences. As justification for this hypothesis, consider the
\texttt{Fibonacci} benchmark. There are no memory accesses to the heap, and the
code leaves little room for optimization (algorithmic rewrites aside), so it is
effectively a benchmark of how fast a runtime can execute fork-join
parallelism.

@@

There are two cases we examine using the profiler. The first is the case when
there is a large difference in run times between implementations. For these our
hypothesis is that the slower implementations are for some reason spending more
time in the runtime. As mentioned, it is difficult to understand exactly why an
implementation is spending more time in the runtime without further
investigation, but it is a useful sanity check nonetheless. For this case, we
turn to the \texttt{Fib} benchmark.  Indeed, we get roughly the following
breakdowns of work-to-overhead ratio according to the Linux \texttt{perf} tool.
\begin{itemize}
\item \texttt{tapir: ~30\% runtime overhead, ~50\% work, ~20\% other}
\item \texttt{clang: ~85\% runtime overhead, ~5\% work, ~10\% other}
\item \texttt{icc: ~85\% runtime overhead, ~5\% work, ~10\% other}
\item \texttt{gcc: ~100\% runtime overhead, ~0\% work, ~0\% other}
\end{itemize}
This supports our hypothesis that much of the slowdown incurred is overhead in
the runtime. For Tapir the overhead is relatively small given the size of the
work chunks, for \texttt{clang} and \texttt{icc} it is significantly larger,
and for \texttt{gcc} it is overwhelming. It's worth noting that \texttt{clang} 
and \texttt{icc} have effectively the same runtime, so the fact that their 
overheads are similar is unsurprising.

@@

Another way we can discriminate the cause of performance discrepency is by disabling 
Tapir optimizations. This allows us to directly measure the impact that Tapir
optimizations are having on runtime. For \texttt{fibonacci}, for example,
disabling optimizations results in a roughly ~2x performance loss. This is
likely due to Tapir inlining the second task call and moving some memory
operations into registers. This is strong evidence that Tapir optimizations are
an important factor in the performance of compiled task-parallel code. Again,
note that this compounds with runtime performance discrepency to result in the
measured performance reported in Section~\ref{Sec:Results}.
